
```{r {{fSetName}}-setup}
sOutDir <- sprintf('%s/%s', outDir, fSetName)
if (!dir.exists(sOutDir)) dir.create(sOutDir)

sDataDir <- sprintf('%s/data', sOutDir)
if (!dir.exists(sDataDir)) dir.create(sDataDir)

sFigDir <- sprintf('%s/figures', sOutDir)
if (!dir.exists(sFigDir)) dir.create(sFigDir)

knitr::opts_chunk$set(fig.path = paste0(sFigDir, '/'))
```

All mentioned result files, both data or figures (if any), from the analysis within this section can be found in folder `r sprintf('*%s*', fSetName)`.

```{r {{fSetName}}_clean-data}
# work with gene name, exclude duplicated gene name by retaining the copy that has the largest 
# number of read counts
dupNames <- unique(expCnt[duplicated(expCnt[, expNameCol]), expNameCol])
if (length(dupNames) > 0) {
  write.table(dupNames, paste0(sDataDir, '/dup-genes.txt'), 
            row.names = FALSE, col.names = FALSE, quote = FALSE)
}

isRowToDel <- rep(FALSE, nrow(expCnt))
for (dupName in dupNames) {
  iRows <- which(expCnt[, expNameCol] == dupName)
  iRowMax <- which.max(rowSums(abs(expCnt[iRows, expDataSCol:ncol(expCnt)])))
  iRowToDel <- iRows[-iRowMax]
  isRowToDel[iRowToDel] <- TRUE
}
expCnt <- expCnt[!isRowToDel, ]
cntMat <- as.matrix(expCnt[, expDataSCol:ncol(expCnt)])
rownames(cntMat) <- expCnt[, expNameCol]

# remove duplicated feature from TF binding as well
featTf <- featTf[expCnt[, expIdCol], ]
rownames(featTf) <- expCnt[, expNameCol]

# use gene name for TF as well
colnames(featTf) <- expCntFull[match(colnames(featTf), expCntFull[, expIdCol]), expNameCol]

# exclude TFs that have no expression
nTtlTf <- ncol(featTf) # the total number TFs in the input featTf matrix
tfAct <- colnames(featTf)[colnames(featTf) %in% rownames(cntMat)]
featTf <- featTf[, tfAct]

sumFltDup <- function() {
  desc <- sprintf('There are %d %ss that have duplicated rows (by name of the %s) ', length(dupNames), feature, feature)
  if (inputParams$coeMode == 'lfc') {
    paste0(desc, 'in the given log of fold change file ')
  } else {
    paste0(desc, 'in the given read count file ')
  }
  desc <- paste0(desc, 'and were saved to *dup-genes.txt*.' )
  desc <- sprintf('%s For each such %s, the duplicate with the maximum ', desc, feature)
  if (inputParams$coeMode == 'lfc') {
    desc <- paste0(desc, 'sum of absolute value of log fold change across all pairwise comparisons ')
  } else {
    desc <- paste0(desc, 'total counts across all samples ')
  }
  desc <- paste0(desc, 'was retained and all the rest were removed. ',
                 sprintf('After the removal, there were %s %ss left for analysis. ', printNum(nrow(featTf)), feature))
  return(desc)
}
```

`r if (length(dupNames) > 0) sumFltDup()` `r if (nTtlTf - ncol(featTf) > 0) sprintf('Among the %d TFs, there are %d TFs that had no expression and were excluded, leaving %d for subsequent analysis. ', nTtlTf, nTtlTf - ncol(featTf), ncol(featTf))`

### Network reconstruction by predicted TF binding only

```{r {{fSetName}}_grn-mot}
# network constructed by considering TF binding
netByTB <- featTf
# exclude self-edge
for (tf in tfAct) {
  netByTB[tf, tf] <- 0
}
write.csv(netByTB, paste0(sDataDir, '/net-by-tb.csv'), quote = FALSE)

# total TF-target pairs
nTtlPairs <- nrow(netByTB) * ncol(netByTB)
```

A network was reconstructed based on only the predicted TF binding (saved to *net-by-tb.csv*). In this network, there are `r printNum(length(which(netByTB == 1)))` edges, i.e., `r sprintf('%.2f%%', length(which(netByTB == 1)) * 100 / nTtlPairs)` out of the total `r printNum(nTtlPairs)` TF-target pairs. 

The number of regulating TFs was counted for each gene. The plot below (*nTF-by-tb-hist-1.pdf*) shows its distribution among all genes. The data used in this plot can be found in *tf-by-tb-stat.csv*.

```{r nTF-by-tb-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTfByTBPlotDf <- data.frame(nTf = apply(netByTB, 1, function(x) return(sum(x))))
write.csv(cbind(gene = rownames(netByTB), nTfByTBPlotDf),
          paste0(sDataDir, '/tf-by-tb-stat.csv'), row.names = F, quote = F)

ggplot(nTfByTBPlotDf, aes(x = nTf)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of regulating TFs', y = 'Frequency of genes')
rm(nTfByTBPlotDf)
```

The number of target genes was counted for each TF. The following histogram (*nTrt-by-tb-hist-1.pdf*) shows its distribution among all TFs. The data used in the plot can be found in *trt-by-tb-stat.csv*.

```{r nTrt-by-tb-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTrtByTBPlotDf <- data.frame(nTrt = apply(netByTB, 2, function(x) return(sum(x))))
write.csv(cbind(colnames(netByTB), nTrtByTBPlotDf),
          paste0(sDataDir, '/trt-by-tb-stat.csv'), row.names = F, quote = F)
ggplot(nTrtByTBPlotDf, aes(x = nTrt)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of targets', y = 'Frequency of TFs')
rm(nTrtByTBPlotDf)
```

### Network reconstruction by gene co-expression only

```{r {{fSetName}}_grn-exp}
# calculate correlation among features
if (inputParams$coeMode == 'group-average') {
  # use group average to calculate correlation
  # load group specification
  sampGrpSpec <- read.csv(inputParams$expSampGrpSpec)
  sampGrps <- unique(sampGrpSpec$Group)
  cntMatTmp <- matrix(rep(0, nrow(cntMat) * length(sampGrps)), nrow = nrow(cntMat))
  rownames(cntMatTmp) <- rownames(cntMat)
  colnames(cntMatTmp) <- sampGrps
  for (grp in sampGrps) {
    cntMatTmp[, grp] <- apply(cntMat[, sampGrpSpec$Sample[sampGrpSpec$Group == grp], drop = F], 1, 
                              function(x) return(mean(x)))
  }
  cntMat <- cntMatTmp
  rm(sampGrpSpec, sampGrps, cntMatTmp)
} 
fcor <- cor(t(cntMat), t(cntMat[tfAct,]), method = 'pearson')
write.csv(data.frame(gene = rownames(cntMat), fcor), paste0(sDataDir, '/gene-coe-cor.csv'),
          row.names = F, quote = F)

netByCe <- matrix(rep(0, nrow(fcor) * ncol(fcor)), nrow = nrow(fcor))
colnames(netByCe) <- colnames(fcor)
rownames(netByCe) <- rownames(fcor)

netByCe[fcor >= minCor] <- 1    # 1 indicating activate
netByCe[fcor <= -1 * minCor] <- -1 # -1 indicating inhibit
# exclude self-edge
for (tf in tfAct) {
  netByCe[tf, tf] <- 0
}
write.csv(netByCe, paste0(sDataDir, '/net-by-ce.csv'), quote = FALSE)

printCoeMode <- function() {
  if (inputParams$coeMode == 'single-gene') {
    print('expression quantification at individual gene level')
  } else if (inputParams$coeMode == 'group-average'){
    print('group average of expression quantification')
  } else if (inputParams$coeMode == 'lfc') {
    print('log of fold change from pairewise comparisons')
  } else {
    print('unknown coeMode mode')
  }
}
```

A network was also reconstructed based on co-expression between the TF and the target `r feature`. The co-expression between each pair of TF and target `r feature` was measured by Pearson correlation calculated using `r printCoeMode()`. The obtained correlation for all pairs can be found in *gene-coe-cor.csv*. An edge between a TF and a target was established when the absolute value of the correlation between them was no less than `r minCor`, with the positive correlation implying positive regulation (i.e., promotion) and negative correlation indicating negative regulation (i.e., inhibition). The reconstructed network was saved to *net-by-ce.csv*, in which value 0 means no edge, 1 indicates positive regulation, -1 represents negative regulation. In this network, there are `r printNum(length(which(netByCe == 1)))` (i.e., `r sprintf('%.2f%%', length(which(netByCe == 1)) * 100 / nTtlPairs)` out of all TF-target pairs) positive regulatory links (edges) and `r printNum(length(which(netByCe == -1)))` (i.e., `r sprintf('%.2f%%', length(which(netByCe == -1)) * 100 / nTtlPairs)` out of all TF-target pairs) negative regulatory links. 

The number of regulating TFs was counted for each gene. The plot below (*nTF-by-ce-hist-1.pdf*) shows its distribution among all genes. The data used in this plot can be found in *tf-by-ce-stat.csv*.

```{r nTF-by-ce-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTfByCePlotDf <- data.frame(nTf = apply(netByCe, 1, function(x) return(sum(abs(x)))))
write.csv(cbind(gene = rownames(netByCe), nTfByCePlotDf),
          paste0(sDataDir, '/tf-by-ce-stat.csv'), row.names = F, quote = F)

ggplot(nTfByCePlotDf, aes(x = nTf)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of regulating TFs', y = 'Frequency of genes')
rm(nTfByCePlotDf)
```

The number of target genes was counted for each TF. The following histogram (*nTrt-by-ce-hist-1.pdf*) shows its distribution among all TFs. The data used in the plot can be found in *trt-by-ce-stat.csv*.

```{r nTrt-by-ce-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTrtByCePlotDf <- data.frame(nTrt = apply(netByCe, 2, function(x) return(sum(abs(x)))))
write.csv(cbind(gene = colnames(netByCe), nTrtByCePlotDf),
          paste0(sDataDir, '/trt-by-ce-stat.csv'), row.names = F, quote = F)
ggplot(nTrtByCePlotDf, aes(x = nTrt)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of targets', y = 'Frequency of TFs')
rm(nTrtByCePlotDf)
```

### Network reconstruction by considering both predicated TF binding and gene co-expression

```{r {{fSetName}}_grn-both} 
netByBoth <- netByCe
netByBoth <- netByBoth[rownames(netByTB), colnames(netByTB)] # make sure rows and column line up
netByBoth[netByTB == 0] <- 0
write.csv(netByBoth, paste0(sDataDir, '/net-by-both.csv'), quote = FALSE)

# create a sif file for visualization
posLnks <- convMatIdx1to2(which(netByBoth == 1), nrow(netByBoth), ncol(netByBoth))
negLnks <- convMatIdx1to2(which(netByBoth == -1), nrow(netByBoth), ncol(netByBoth))
sif <- rbind(data.frame(tf = colnames(netByBoth)[posLnks[, 2]], type = rep('+', nrow(posLnks)), target = rownames(netByBoth)[posLnks[, 1]]),
             data.frame(tf = colnames(netByBoth)[negLnks[, 2]], type = rep('-', nrow(negLnks)), target = rownames(netByBoth)[negLnks[, 1]]))
# sort by TF
sif <- sif[sort(sif$tf, index.return = TRUE)$ix, ]
write.table(sif, paste0(sDataDir, '/net-by-both.sif'), quote = FALSE, 
            row.names = FALSE, col.names = FALSE, sep = '\t')
```

The above two networks (i.e., one by predicted TF binding only and one by co-expression only) were combined to reconstruct the final network by removing links in the network by co-expression that do not appear in the network by predicted TF binding. The final network was saved to *net-by-both.csv*. In this network, there are `r printNum(length(which(netByBoth == 1)))` (i.e., `r sprintf('%.2f%%', length(which(netByBoth == 1)) * 100 / nTtlPairs)` out of all pairs) positive regulatory links (edges) and `r printNum(length(which(netByBoth == -1)))` (i.e., `r sprintf('%.2f%%', length(which(netByBoth == -1)) * 100 / nTtlPairs)` out of all pairs) negative regulatory links. 

The number of regulating TFs was counted for each gene. The plot below (*nTF-by-both-hist-1.pdf*) shows its distribution among all genes. The data used in this plot can be found in *tf-by-both-stat.csv*.

```{r nTF-by-both-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTfByBothPlotDf <- data.frame(nTf = apply(netByBoth, 1, function(x) return(sum(abs(x)))))
write.csv(cbind(gene = rownames(netByBoth), nTfByBothPlotDf),
          paste0(sDataDir, '/tf-by-both-stat.csv'), row.names = F, quote = F)

ggplot(nTfByBothPlotDf, aes(x = nTf)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of regulating TFs', y = 'Frequency of genes')
rm(nTfByBothPlotDf)
```

The number of target genes was counted for each TF. The following histogram (*nTrt-by-both-hist-1.pdf*) shows its distribution among all TFs. The data used in the plot can be found in *trt-by-both-stat.csv*.

```{r nTrt-by-both-hist, warning=FALSE, fig.height=4, fig.width=6.5, dev=c('png', 'pdf')}
nTrtByBothPlotDf <- data.frame(nTrt = apply(netByBoth, 2, function(x) return(sum(abs(x)))))
write.csv(cbind(gene = colnames(netByBoth), nTrtByBothPlotDf),
          paste0(sDataDir, '/trt-by-both-stat.csv'), row.names = F, quote = F)
ggplot(nTrtByBothPlotDf, aes(x = nTrt)) + 
  geom_histogram(color="darkblue", fill = "lightblue", bins = 100) + 
  theme_classic()  +
  theme(axis.text.x = element_text(size = 12), 
        axis.title.x = element_text(size = 12), axis.title.y = element_text(size = 12)) +
  labs(x = '# of targets', y = 'Frequency of TFs')
rm(nTrtByBothPlotDf)
```

```{r grn-clean-up}
knitr::opts_chunk$set(fig.path = paste0(figDir, '/'))
```